---
title: 'Project: Risk of Presence of Cardiovascular Disease'
author: "Waner Li 873464868"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message = F, warning = F}
library(tidyverse)
library(broom)
library(glmnet)
library(caret)
library(ISLR)
library(janitor)
library(plotROC)
library(kernlab)
library(stringr)
library(rpart)
library(rpart.plot)
library(partykit)
library(MASS)
library(randomForest)
library(tree)
library(gbm)
library(ranger)
library(dplyr)
library(GGally)
library(corrplot)
library(MASS)
library(pROC)
library('scales')
theme_set(theme_bw())
```

```{r}
set.seed(1982)
df<-read.csv("cardio_train.csv")
ori<-read.csv("cardio_train.csv")

# clean age
df <- df %>%
  mutate(age = age/365) %>%
  mutate(BMI = weight/((height/100)^2))
summary(df)
summary(ori)

```

```{r}
# clean height outliers as min height was 55cm and max height was 250cm which are most likely not correct
clean_height <- df$height <= as.numeric(quantile(df$height, probs = c(0.9999))) & df$height >= as.numeric(quantile(df$height, probs = c(0.0001)))

# clean weight outliers
clean_weight <- df$weight <= as.numeric(quantile(df$weight, probs = c(0.9999))) & df$weight >= as.numeric(quantile(df$weight, probs = c(0.0001)))

# clean ap_hi outliers
clean_ap_hi <- df$ap_hi <= as.numeric(quantile(df$ap_hi, probs = c(0.999))) & df$ap_hi >= as.numeric(quantile(df$ap_hi, probs = c(0.0001)))

# clean ap_lo outliers
clean_ap_lo <- df$ap_lo <= as.numeric(quantile(df$ap_lo, probs = c(0.9999))) & df$ap_lo >= as.numeric(quantile(df$ap_lo, probs = c(0.0001)))


# ap_hi >= ap_lo
clean_bp <- df$ap_hi >= df$ap_lo

# clean BMI outliers


clean_bmi <- df$BMI <= 60 & df$BMI >= 10

clean_vec <- clean_height&clean_weight&clean_ap_lo&clean_ap_hi&clean_bp&clean_bmi
df_cardio <- df[which(clean_vec),]

# use BMI to replace height and weight
df <- df_cardio %>%
  dplyr::select(-id,-height,-weight) %>%
  na.omit()



summary(df)
str(df)
```

Multiple logistic regression model

Deselect gender as its p value is low
```{r,include=FALSE}
full_glm <- glm(cardio ~ .,
                family = "binomial", 
                data = df)
summary(full_glm)
df <- df %>%
  dplyr::select(-gender)
```

Multivariate analysis - Correlation 
```{r}
ggcorr(df, method = c("everything", "pearson")) 
```

```{r}
abs(cor(df))>0.7
```


### The Relation between the Systolic blood pressure and Diastolic blood pressure

The features are not all independent with each other. Take the relation betweeen the Systolic blood pressure and Diastolica=ggplot(data,aes(x=age,y=target))+geom_point()+geom_smooth(color="yellow")
b=a+scale_x_continuous(name="Age")+scale_y_continuous(name="Target")
b+ggtitle("Scatter plot with smooth curve between age and target") blood pressure for example. So we should deal with the coefficient problem. An idea to achieve this is using lasso.
```{r}
ggplot(data = df,aes(x=ap_hi,y=ap_lo)) + geom_point()
```

Age
```{r}
str(df)
ageplot <- ggplot(df,aes(x = age,y = cardio))
ageplot + geom_point(col = "#6e0000", alpha = .25) +
  geom_smooth(color="yellow",size = 2) +
  scale_x_continuous(name="Age") + 
  scale_y_continuous(name="Cardio") +
  ggtitle("Cardio vs age") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
glm_cardio <- glm(cardio ~ poly(BMI,2), data = df, family = "binomial")
tidy(glm_cardio)
```

Splines
```{r}
ss_fits <- smooth.spline(df$BMI, df$cardio)
df_ss <- tibble(x = ss_fits$x, y = ss_fits$y)

bmi_ss <- ggplot(data = df,
                 aes(x = BMI, y = cardio))

bmi_ss + geom_point(alpha = 0.25, color = "#6e0000", alpha = .25) +
  geom_line(data = df_ss,
            aes(x = x, y = y),
            color="yellow",
            size = 2)  +
  scale_x_continuous(name="BMI") + 
  scale_y_continuous(name="Cardio") +
  ggtitle("Cardio vs BMI") +
  theme(plot.title = element_text(hjust = 0.5))
```

Cholesterol and Glucose
```{r}
# show the distribution of cholesterol

df_dis <- filter(df,cardio == "1")
ggplot(df_dis, aes(x = cholesterol)) +
        geom_histogram(aes(y = ..density..), binwidth = 0.5)

# show the distribution of gluc
ggplot(df, aes(x = gluc)) +
        geom_histogram(aes(y = ..density..), binwidth = 0.5)
```

Smoking, Alcohol intake, Physical activity

```{r}

# show the distribution of smoke
ggplot(df, aes(x = smoke)) +
        geom_histogram(aes(y = ..density..), binwidth = 0.5)

# show the distribution of alco
ggplot(df, aes(x = alco)) +
        geom_histogram(aes(y = ..density..), binwidth = 0.5)

# show the distribution of active
ggplot(df, aes(x = active)) +
        geom_histogram(aes(y = ..density..), binwidth = 0.5)

```




LDA
```{r}
set.seed(12345)
df$cardio <- as.factor(df$cardio)
inTraining <- caret::createDataPartition(df$cardio, 
                                         p = .75,
                                         list = F)
training <- df[inTraining, ]
testing  <- df[-inTraining, ]
```

```{r}
set.seed(12345)
cardio_lda <- lda(cardio ~ ., data = training)
```

```{r}
set.seed(12345)
acc_lda <- confusionMatrix(table(predict(cardio_lda, newdata = testing)$class, testing$cardio), positive = "1")
pct_lda <- unname(acc_lda$overall["Accuracy"]) * 100
pct_lda
```

Fit a regression tree to the training set. Plot the tree, and interpret
the results. What test error rate do you obtain?

```{r}
set.seed(12345)
test <- df %>%
  mutate(cardio = as.numeric(cardio)) %>%
  mutate(cardio = if_else(cardio == "1",0,1))
str(test)

inTraining <- caret::createDataPartition(test$cardio, 
                                         p = .5,
                                         list = F)
f_training <- test[inTraining, ]
f_testing  <- test[-inTraining, ]

cardio_tree <- rpart::rpart(cardio ~ .,data = f_training)
prp(cardio_tree)
```


```{r}
cardio_tree
```

```{r}
set.seed(12345)
tree_test_preds<-predict(cardio_tree, newdata = f_testing)

tree_accuracy <- mean(f_testing$cardio == round(tree_test_preds, digits = 0)) * 100
```


Bagging
```{r}
set.seed(12345)
inTraining <- caret::createDataPartition(df$cardio, 
                                         p = .75,
                                         list = F)
training <- df[inTraining, ]
testing  <- df[-inTraining, ]

```

```{r}
set.seed(12345)
bag_cardio <- ranger::ranger(cardio ~., 
                             data = training, 
                             mtry = 9,
                             probability = TRUE)
bag_cardio

bag_test_preds<-predict(bag_cardio, data = testing)

bag_accuracy <- mean(testing$cardio == round(bag_test_preds$predictions[,2], digits = 0)) *100
bag_accuracy
```


Random Forest 
```{r}
set.seed(12345)

tune_grid<-expand.grid(mtry = 2:9,
                       splitrule = "gini",
                       min.node.size = 10)

train_control<-trainControl(method = "cv", number = 10)

rf_cardio_cv <- train(cardio ~.,
                      data = training,
                      method = "ranger",
                      num.trees = 15,
                      importance = "impurity",
                      trControl = train_control,
                      tuneGrid = tune_grid)

rf_cardio_cv
```



```{r}
plot(rf_cardio_cv)
```
Best mtry at each step is yielding the lowest RMSE




```{r, include=FALSE}
rf_cardio_8 <- ranger(cardio~.,
                      data = training,
                      mtry = ,
                      probability = TRUE)

rf_cardio_8
```


```{r, include=FALSE}
test_preds<-predict(rf_cardio_8, data = testing)

rf_accuracy <- mean(testing$cardio == round(test_preds$predictions[,2], digits = 0)) * 100
```

Importance
```{r, include=FALSE}
imp<-varImp(rf_cardio_cv)$importance
rn<-row.names(imp)
imp_df<-data_frame(variable = rn,
                   importance = imp$Overall) %>%
  arrange(desc(-importance)) %>%
  mutate(variable = factor(variable,variable))
```

```{r}
rf_imp_graph <- ggplot(data = imp_df,
            aes(variable,importance))

rf_imp_graph + geom_col(fill = "#6e0000") + 
  coord_flip()
```



ROC

```{r, include=FALSE}
cardio_lda <- lda(cardio ~ ., data = testing)
fits <- predict(cardio_lda)
new_fits <- mutate(testing, 
                   pprobs = predict(cardio_lda)$posterior[, 2],
                   default = if_else(cardio == "1",1,0))

summary_pred <- new_fits %>%
  mutate(bg_probs = bag_test_preds$predictions[,2])%>%
  mutate(rf_probs = test_preds$predictions[,2])%>%
  dplyr::select(default,pprobs,bg_probs,rf_probs) %>%
  gather("method","prob",-1)
```

```{r}
roc_graph <- ggplot(data = summary_pred,
            aes(d = default, m = prob, col = method))

roc_graph + geom_roc(n.cuts = 0) +
  style_roc() +
  scale_color_brewer(palette = "Dark2")
```



